Manuscript progress report

### 2023-12-04 Docs/ProgressReportAppl.Rmd
WIT presentation figures

# code
source(here("Code/GLMM-FPCA.R")) 
# use pred_latent function to estimate latent function 
source(here("Code/OutsampBayes.R"))
```

@@

@@

```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))

plot(mid, fpca_mod$efunctions[, 1], pch=16, xlab="bin", ylab="PC1", col = "blue",
     cex = 0.5)
lines(1:J, df_phi[, 1])

plot(mid, fpca_mod$efunctions[, 2], pch=16, xlab="bin", ylab="PC2", col = "blue",
     cex = 0.5)
lines(1:J, df_phi[, 2])

plot(mid, fpca_mod$efunctions[, 3], pch=16, xlab="bin", ylab="PC3", col = "blue",
     cex = 0.5)
lines(1:J, df_phi[, 3])

plot(mid, fpca_mod$efunctions[, 4], pch=16, xlab="bin", ylab="PC4", col = "blue",
     cex = 0.5)
lines(1:J, df_phi[, 4])
```

>>>>

# code
source(here("Code/GLMM-FPCA.R")) 
# use pred_latent function to estimate latent function 
K <- 4
source(here("Code/OutsampBayes.R"))
```

@@

```{r PC}
data.frame(t = mid, fpca_mod$efunctions[, 1:4]) %>%
  rename(PC1 = X1, PC2=X2, PC3=X3,PX4=X4) %>%
  pivot_longer(2:5, names_to = "PC") %>%
  ggplot()+
  geom_line(aes(x=t, y=value, col = PC))+
  labs(x="Time", y="", title = "Eigenfunctions on the binned grid")
```

@@

```{r}
df_pc1 <- data.frame(t=1:J,  df_phi) %>%
  rename(PC1=X1, PC2=X2, PC3=X3, PC4=X4) %>%
  pivot_longer(2:5, names_to = "PC")

df_pc2 <- data.frame(t = mid, fpca_mod$efunctions[, 1:4]) %>%
  rename(PC1=X1, PC2=X2, PC3=X3, PC4=X4) %>%
  pivot_longer(2:5, names_to = "PC")

left_join(df_pc1, df_pc2, by = c("t", "PC")) %>%
  ggplot()+
  geom_line(aes(x=t, y=value.x, col = PC), linewidth = 0.5)+
  geom_point(aes(x=t, y=value.y, col = PC), na.rm = T, alpha = 0.5, size = 0.5)+
  labs(x="Time", y="", title = "Re-evaluate eigenfunctions on the original grid")

### 2023-12-04 Docs/manuscript.Rmd
WIT presentation figures

---
title: "Manuscript progress report"
author: "Ying Jin"
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: no
    number_sections: true
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    font: 12pt
    keep_tex: true
---

@@

```{r fig_sim_data}
rand_id <- sample(501:600, size = 4)
pred_list_all[[1]] %>% filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  facet_wrap(~id)+
  labs(title = "Simulated data")
```

@@

```{r fig_pred}
# range(pred_list_all[[1]]$pred0.2, na.rm = T)
fig_fgfpca <- pred_list_all[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA", col = "Maximum observation time")

fig_adglmm <- pred_list_ref[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive", col = "Maximum observation time")

@@

```{r}
ggarrange(fig_fgfpca, fig_adglmm,
          nrow = 1, common.legend = T)
```

@@

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_ise, mean_ise_ref, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Integrated squared error") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

@@

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_auc, mean_auc_ref, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Area under the ROC curve") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

```{r time}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive"),
  "Fit" = c(mean(fit_time)/60, mean(fit_time_ref)),
           "Prediction"= c(mean(pred_time), mean(pred_time_ref)/60)) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Computation time (minutes)") %>%
  kable_styling(full_width = F) 
```

@@

```{r}
data.frame(Window = c("(0.35,0.45]", "(0.45,0.55]", "(0.55,0.65]", "(0.65, 0.75]"),
           ise_mat_subset, ise_mat_subset2, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Integrated squared error") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Observation track" = 8))
```

@@

```{r}
data.frame(Window = c("(0.35,0.45]", "(0.45,0.55]", "(0.55,0.65]", "(0.65, 0.75]"),
           auc_mat_subset, auc_mat_subset2, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Integrated squared error") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Observation track" = 8))
```

@@

```{r nhance example}
df_nhanes %>% select(SEQN, Z, sind) %>% 
  mutate(Z = factor(Z, levels = 0:1, labels = c("Inactive", "Active"))) %>%
  mutate(SEQN = factor(SEQN, labels = 1:8763)) %>%
  ggplot()+
  geom_tile(aes(x=sind, y = SEQN, fill = Z))+
  labs(x="Time", y="Subject", title = "Overview of NHANES binary activity indicator")+
  theme(axis.text.y = element_blank())

@@

colnames(auc_mat_nhanes) <- c("0-8am", "0am-4pm")
```

@@

colnames(auc_mat_nhanes2) <- c("0-8am", "0am-4pm")
```

```{r}
data.frame(Window = c("8am-4pm", "4am-12pm"),
           auc_mat_nhanes, auc_mat_nhanes2, check.names = F) %>%
  kable(digits = 3, table.attr = "style = \"color: black;\"", caption = "Area Under the ROC curve") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 2, "GLMMadaptive" = 2)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 4))
```

@@

```{r}
rand_id2 <- sample(unique(nhanes_pred_fgfpca$id), 4)

ggarrange(
  nhanes_pred_fgfpca %>% 
    filter(id %in% rand_id2) %>%
    mutate_at(vars(starts_with("pred")), function(x)exp(x)/(1+exp(x))) %>% 
    ggplot()+
    geom_line(aes(x=t, y = pred480, col = "8am"), linetype = "dashed", na.rm = T)+
    geom_line(aes(x=t, y = pred960, col = "4pm"),linetype = "dashed", na.rm = T)+
    geom_point(aes(x=t, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "fGFPCA"),

  nhanes_pred_adglmm %>% 
    filter(id %in% rand_id2) %>%
    mutate_at(vars(starts_with("pred")), function(x)exp(x)/(1+exp(x))) %>% 
    ggplot()+
    geom_line(aes(x=t, y = pred480, col = "8am"), linetype = "dashed", na.rm = T)+
    geom_line(aes(x=t, y = pred960, col = "4pm"),linetype = "dashed", na.rm = T)+
    geom_point(aes(x=t, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "GLMMadaptive"),
  nrow = 1, common.legend = T)
```

>>>>

---
title: "Manuscript progress report"
author: "Ying Jin"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
    keep_tex: yes
  html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float: yes
    font: 12pt
    keep_tex: yes
---

@@

```{r fig_sim_data}
rand_id <- sample(501:600, size = 4)
pred_list_all[[1]] %>% filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), show.legend = F)+
  facet_wrap(~id)+
  labs(title = "Simulated data")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))
```

@@

```{r}
# range(pred_list_all[[1]]$pred0.2, na.rm = T)
fig_fgfpca <- pred_list_all[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "fGFPCA", col = "Maximum observation time")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5))

fig_adglmm <- pred_list_ref[[1]] %>%
  filter(id %in% rand_id) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), function(x){exp(x)/(1+exp(x))}) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"))+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linetype="dashed", na.rm=T)+
  facet_wrap(~id)+
  labs(title = "GLMMadaptive", col = "Maximum observation time")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(axis.text = element_text(size = 5))

@@

```{r fig_sim, fig.height=4, fig.width=8}
ggarrange(fig_fgfpca, fig_adglmm,
          nrow = 1, common.legend = T)
```

@@

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_ise, mean_ise_ref, check.names = F) %>%
  kable(digits = 3, caption = "Integrated squared error", booktabs=T) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

@@

```{r}
data.frame(Window = c("(0.2, 0.4]", "(0.4, 0.6]", "(0.6, 0.8]", "(0.8, 1.0]"),
           mean_auc, mean_auc_ref, check.names = F) %>%
  kable(digits = 3, caption = "Area under the ROC curve", booktabs = T) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 8))
```

```{r time}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive"),
  "Fit" = c(mean(fit_time)/60, mean(fit_time_ref)),
           "Prediction"= c(mean(pred_time), mean(pred_time_ref)/60)) %>%
  kable(digits = 3, caption = "Computation time (minutes)", booktabs = T) %>%
  kable_styling(full_width = F) 
```

@@

```{r}
data.frame(Window = c("(0.35,0.45]", "(0.45,0.55]", "(0.55,0.65]", "(0.65, 0.75]"),
           ise_mat_subset, ise_mat_subset2, check.names = F) %>%
  kable(digits = 3, caption = "Integrated squared error") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Observation track" = 8))
```

@@

```{r}
data.frame(Window = c("(0.35,0.45]", "(0.45,0.55]", "(0.55,0.65]", "(0.65, 0.75]"),
           auc_mat_subset, auc_mat_subset2, check.names = F) %>%
  kable(digits = 3, caption = "Integrated squared error") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" "= 1, "Observation track" = 8))
```

@@

```{r nhance_example}
df_nhanes %>% select(SEQN, Z, sind) %>% 
  filter(SEQN %in% unique(nhanes_pred_adglmm$id) | SEQN %in% unique(nhanes_pred_fgfpca$id)) %>%
  mutate(Z = factor(Z, levels = 0:1, labels = c("Inactive", "Active"))) %>%
  mutate(SEQN = factor(SEQN)) %>%
  ggplot()+
  geom_tile(aes(x=sind, y = SEQN, fill = Z))+
  labs(x="Time", y="Subject", title = "Overview of NHANES binary activity indicator")+
  theme(axis.text.y = element_blank())+
  scale_x_continuous(breaks = seq(0, 1440, by = 480),
                     labels = c("Midnight", "8am", "4pm", "Midnight"))
```

@@

colnames(auc_mat_nhanes) <- c("8am", "4pm")
```

@@

colnames(auc_mat_nhanes2) <- c("8am", "4pm")
```

```{r}
data.frame(Window = c("8am-4pm", "4am-midnight"),
           auc_mat_nhanes, auc_mat_nhanes2, check.names = F) %>%
  kable(digits = 3, caption = "Area Under the ROC curve", booktabs = T) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 2, "GLMMadaptive" = 2)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 4))
```

@@

```{r pred_nhanes, fig.height=4, fig.width=10}
rand_id2 <- sample(unique(nhanes_pred_fgfpca$id), 4)

ggarrange(
  nhanes_pred_fgfpca %>% 
    filter(id %in% rand_id2) %>%
    mutate_at(vars(starts_with("pred")), function(x)exp(x)/(1+exp(x))) %>% 
    ggplot()+
    geom_line(aes(x=t, y = pred480, col = "8am"), linetype = "dashed", na.rm = T)+
    geom_line(aes(x=t, y = pred960, col = "4pm"),linetype = "dashed", na.rm = T)+
    geom_point(aes(x=t, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "fGFPCA")+
  scale_x_continuous(breaks = seq(0, 1440, by = 480), 
                     labels = c("Midnight", "8am", "4pm", "Midnight"))+
  theme(axis.text = element_text(size = 5)),

  nhanes_pred_adglmm %>% 
    filter(id %in% rand_id2) %>%
    mutate_at(vars(starts_with("pred")), function(x)exp(x)/(1+exp(x))) %>% 
    ggplot()+
    geom_line(aes(x=t, y = pred480, col = "8am"), linetype = "dashed", na.rm = T)+
    geom_line(aes(x=t, y = pred960, col = "4pm"),linetype = "dashed", na.rm = T)+
    geom_point(aes(x=t, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "GLMMadaptive")+
  scale_x_continuous(breaks = seq(0, 1440, by = 480), 
                     labels = c("Midnight", "8am", "4pm", "Midnight"))+
  theme(axis.text = element_text(size = 5)),
  nrow = 1, common.legend = T)
```

```{r small_exp}
# a smaller example of binary functions
df_nhanes %>% select(SEQN, Z, sind) %>% 
  filter(SEQN %in% rand_id2 & Z == 1) %>% 
  mutate(SEQN = as.factor(SEQN)) %>% 
  ggplot()+
  geom_point(aes(x=sind, y = SEQN), size = 0.5)+
  labs(x="Time", y="Subject", title = "NHANES binary activity indicator")
```

# Additional figures

## fGFPCA algorithm demonstration

```{r binning}
df_nhanes %>% select(SEQN, Z, sind) %>% 
  filter(SEQN %in% rand_id2 & sind >= 650 & sind <= 750) %>% 
  mutate(SEQN = as.factor(SEQN)) %>% 
  ggplot()+
  geom_point(aes(x=sind, y = Z, col = SEQN), size = 0.5)+
  geom_vline(xintercept = seq(650, 750, by = 10), linetype="dashed")+
  scale_y_continuous(breaks = c(0, 1))+
  labs(x="Time", y="", title = "NHANES binary activity indicator")
```

```{r}
df_fig <- df_nhanes %>% select(SEQN, Z, sind) %>% 
  filter(SEQN %in% rand_id2 & sind >= 650 & sind <= 750) %>% 
  mutate(bin = cut(sind, breaks = seq(650, 750, by = 10), 
                   labels = seq(655, 745, by = 10),
                   include.lowest = T)) 
bin_mid <- seq(655, 745, by = 10)

for(i in bin_mid){
  df_bin_i <- df_fig %>% filter(bin==i) 
  glm_i <- glmer(Z ~ 1 + (1|SEQN), data = df_bin_i, family = binomial, nAGQ=0)
  eta_i <- predict(glm_i, type = "link")

  df_fig$eta_hat[df_fig$bin==i] <- eta_i
}

df_fig %>% 
  mutate(SEQN = as.factor(SEQN),
         eta_hat = exp(eta_hat)/(1+exp(eta_hat)),
         bin=as.numeric(as.character(bin))) %>%

  ggplot()+
  geom_point(aes(x=sind, y = Z, col = SEQN), size = 0.5)+
  geom_vline(xintercept = seq(650, 750, by = 10), linetype="dashed")+
  geom_point(aes(x=bin, y=eta_hat, col = SEQN))+
  geom_line(aes(x=bin, y=eta_hat, col = SEQN))+
  scale_y_continuous(breaks = c(0, 1))+
  labs(x="Time", y="", title = "NHANES binary activity indicator")
